---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(hlatools)
```

# hlatools

A collection of tools to work with IPD-IMGT/HLA data

## Main functionality

### Fetch HLA data

A number of functions exist that grab HLA-related data from the internet

* `fetch_IMGTHLA`: Clone the _ANHIG/IMGTHLA_ repo on GitHub to a location specified by the option `hlatools.local_repos`. Defaults to _~/local/db/_.

* `update_IMGTHLS`: Send a pull request to _ANHIG/IMGTHLA_.

* `read_hla_xml(remote = FALSE)`: Read the _hla.xml_ file either fetching it from the IMGT/HLA ftp server or the local _ANHIG/IMGTHLA_ clone.

### Parse HLA data

A HLA locus can be read into a `HLAAllele` object:

```{r}
doc  <- hlatools::read_hla_xml()
dpb1 <- hlatools::parse_hla_alleles(doc, "DPB1")
dpb1
```

A number of accessor functions can be used to work with these objects:

```{r}
dpb1_cmpl <- dpb1[hlatools::is_complete(dpb1)]
hlatools::allele_name(dpb1_cmpl)
hlatools::cwd_status(dpb1_cmpl)
hlatools::sequences(dpb1_cmpl)
```

At a slightly higher level a `HLAAllele` can be encapsulated in an `R6`-based `HLAGene` object:

```{r}
x <- hlatools::HLAGene("DPB1", db_version = "3.28.0")
x

hlatools::db_version(x)
hlatools::locusname(x)
x[hlatools::cwd_status(x) == "Common"]
```

These obects come with the same set of getter functions as `HLAAllele`s plus an additional set of `R6`-methods that implement an API used mostly in conjunction with the `DR2S` package:

* `x$get_closest_complete_neighbor(allele)`: Find the closest full-length allele based on the genetic distance between exon 2 sequences.

* `x$get_reference_sequence(allele)`: Return a `BStringSet` object of a full-length reference sequence for `allele`. If `allele` is not fully known the missing stretches are taken from the allele returned by a call to `get_closest_complete_neighbor()`.

